---
title: "Getting & Cleaning Data: Week 4 Project Codebook"
author: "J. Gomez"
date: "9/12/2019"
output: pdf_document
---

## Codebook

This codebook corresponds to Coursera class, Getting & Cleaning Data, week 4 project assignment.
The scripts corresponding to this project can be found at <https://github.com/reivaj23/Course3Project>

The dataset was obtained from <https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>.

#  Experiment Description

Per the UCI website, the experiments were carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, 3-axial linear acceleration and 3-axial angular velocity were captured at a constant rate of 50Hz.

A complete description of the experiment can be found at <http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones>

# Feature Selection

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

tBodyAcc-XYZ
tGravityAcc-XYZ
tBodyAccJerk-XYZ
tBodyGyro-XYZ
tBodyGyroJerk-XYZ
tBodyAccMag
tGravityAccMag
tBodyAccJerkMag
tBodyGyroMag
tBodyGyroJerkMag
fBodyAcc-XYZ
fBodyAccJerk-XYZ
fBodyGyro-XYZ
fBodyAccMag
fBodyAccJerkMag
fBodyGyroMag
fBodyGyroJerkMag

The data from the acceleration signal from the smartphone accelerometer is recorded in standard gravity units 'g'.
The data from the angular velocity vector measured by the gyroscope are in radians/second. 

# Data Cleaning

The following activities are included in the dataset for this assignmenet are:

```{r setup, include=FALSE}
## Load libraries
library(dplyr)

mydata <- read.csv(file = "mydata.txt", header = TRUE, sep="")
```

```{r, echo=FALSE}
mydata %>% 
    select(activity, activityDesc) %>% 
    unique() %>% 
    knitr::kable(align = 'c', col.names=c("Activity", "Description"))
```

The set of variables that were estimated from these signals for this assignment are: 

mean(): Mean value
std(): Standard deviation

The table below shows the list of measurements included in the dataset. Mean values were computed for each measurement, grouped by activity for each volunteer.

```{r, echo=FALSE}
mydata %>% 
    select(-(subject:activityDesc)) %>% 
    colnames() %>% 
    knitr::kable(align = 'c', col.names="Measurements", row.names = T)
```
