---
title: "Getting & Cleaning Data: Week 4 Project Codebook"
author: "J. Gomez"
date: "9/12/2019"
output: pdf_document
---

## Codebook

This codebook corresponds to Coursera class, Getting & Cleaning Data, week 4 project assignment.
The scripts corresponding to this project can be found at <https://github.com/reivaj23/Course3Project>

The dataset was obtained from <https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>.

#  Experiment Description

The experiments were carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

# Dataset description

The following activities are included in the dataset:

```{r setup, echo=FALSE}
## Load libraries
library(dplyr)

activitylabels <- read.csv("./UCI HAR Dataset/activity_labels.txt", header = FALSE, sep = "", col.names = c("activity", "activityDesc"))
knitr::kable(activitylabels, align = 'c', col.names=c("Activity", "Description"))
```

```{r, echo=FALSE}
features <- read.csv("./UCI HAR Dataset/features.txt", header = FALSE, sep = "")
mean_cols <- grep("mean()", colnames(features), fixed = T)
std_cols <- grep("std()", colnames(features), fixed = T)
knitr::kable(features[2][sort(c(mean_cols, std_cols)),], align = 'c', col.names = "Measurement")
```

